## Jeffrey Dean

## Data Wrangling Final Project-Code Part 1 of 2

## Bank Products Issues, Fraud, and the People Affected
```{r}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
library("janitor")
library("vtree")
library("tidyverse")
library("dplyr") 
library("tidytext") 
library("stringr")
library("rvest")
library("stopwords")
library("choroplethrMaps")
library("choroplethr")
library("plyr")
library("lubridate")
library("clean")
library("cleaner")
library("ggQC")
```

## Code will support what banks have officially been served complaints by their users and how responsive they are to such issues.  In addition the work will support which institutions should improve and in what areas.  As a follow up I will pull information on identity theft of personal information and how this impacts similar sets of customers.  

## To compare customers with banks, we must first build profiles on which users are utilizing credit products and how it is used.  From there I will look at patterns and trends.  

## Section 1 

## First data set is customer application dataset for credit products.
```{r}
customer_data <- read.csv(file = 'customer_application_data.csv')
unique_customers <- length(unique(customer_data$SK_ID_CURR))
unique_customers
```

## The customer count above supports that there is 307k unique values, so this is a large normalized data set.  
```{r}
customer_data_gender <- ggplot(customer_data, aes(x=factor(CODE_GENDER))) +
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  theme_minimal()
customer_data_gender 
```

## The above shows a graphic that has almost twice the number of female credit customers versus male customers. This graph and the below supports a profile of active users of credit in a normal banking institution. 
```{r}
customer_data_education_type <- ggplot(customer_data, aes(x=factor(NAME_EDUCATION_TYPE))) +
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
customer_data_education_type
```

## The education type of a typical customer is secondary or secondary special.  This makes sense as many people have some time of higher level degree or secondary education completed, especially if they have the income to take out credit products. 
```{r}
customer_data_income_type <- ggplot(customer_data, aes(x=factor(NAME_INCOME_TYPE))) +
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
customer_data_income_type
```

## Most credit consumers are working for salary or potentially a business owner as the typical age is working age, with some people being older taking out a loan on a house or other property. 
```{r}
customer_data_family_status <- ggplot(customer_data, aes(x=factor(NAME_FAMILY_STATUS))) +
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  theme_minimal()
customer_data_family_status
```

## Most data in the above graph supports our customers are married, with some being separated or single.  Married age people are ones looking for property or looking for a business loan to implement for a business.  
```{r}

tabyl(customer_data, CODE_GENDER, NAME_CONTRACT_TYPE)
tabyl(customer_data, CODE_GENDER, NAME_FAMILY_STATUS)
tabyl(customer_data, CODE_GENDER, NAME_EDUCATION_TYPE)
tabyl(customer_data, NAME_HOUSING_TYPE, FLAG_OWN_REALTY)
tabyl(customer_data, NAME_EDUCATION_TYPE, OCCUPATION_TYPE)
tabyl(customer_data, NAME_INCOME_TYPE, NAME_EDUCATION_TYPE)
```

## Above we see similiar data as the prior but we can see two variable matrix comparisons for each groupings.  The gender versus family show the person of interest, while items such as housing and if the realty is owned by the individual shows the landscape of the market.  Housing and co-ops are overwhelmingly owned outright while other property otherwise is split because other real estate can be controlled by a management firm.
```{r}
vtree(customer_data, c("NAME_HOUSING_TYPE", "FLAG_OWN_REALTY"), 
      fillcolor = c( LanguageGroup = "#e7d4e8", Gender = "#99d8c9"),
      horiz = FALSE)
```

## The last two graphs support a tree model for the two variables similiar to the profile data in the earlier graphs.  These graphs better support split scenarios as a treat to show the weights of where the data fits.  The tables above show the housing market from a different perspective from the last set, while the income and education shows that housing and investment clearly skews to higher incomes and higher education levels. 
```{r}
customer_data_income_cleaned <- customer_data %>% filter(AMT_INCOME_TOTAL < 5000000)
income_vs_credit_applied <- ggplot(customer_data_income_cleaned, aes(x=AMT_INCOME_TOTAL, y=AMT_CREDIT)) + geom_point() + scale_y_continuous(label=scales::comma) +
scale_x_continuous(label=scales::comma) + 
ggtitle("Income vs Credit Requested")
income_vs_credit_applied
```

## Above shows a cleaned data set and filtered for outliers in order to see patterns in the plot used.  While there are a few outliers with high income, these could either be potential mistakes or wealthy individuals.  This is purely to make the graph readable to users.  There is a loose positive relationship betweeen credit and income levels. Some high income levels proposed to take out middle level of credit, while some lower income individuals have taken a very high level of credit used.  Possibly for a business or for real estate. 

## Section 2

## Bank Complaint Data and Credit Institutions
```{r}
bank_complaint_data <- read.csv(file = 'Bank_Complaint_Data.csv')

tabyl(bank_complaint_data, Timely.response., Company.response.to.consumer)
```

## The data above support a relationship between the status of a complaint and how timely the response was filed.  This is a good starting place to see if complaints as a whole are being handled by institutions.  The table supports that a vast majority of cases (1 million +) are handled and resolved. Only 4143 cases were not timely, maybe because the issues took longer than expected when the cases were filed. The cases that were untimely is less than 1% and it seems to be insignificant in the grand scheme of the cases. 
```{r}
bank_complaint_data_response_time <- ggplot(bank_complaint_data, aes(x=factor(Timely.response.))) +
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  theme_minimal()
bank_complaint_data_response_time
```

## The ggplot for response time supports a graph of timely responses to complaint cases.  The chart supports "yes" as the response for an overwhelming number of cases.
```{r}
bank_complaint_issues <- count(bank_complaint_data, "Issue")
ranked_bank_complaint_issues <- bank_complaint_issues %>% arrange(desc(freq))
head(ranked_bank_complaint_issues, 30)

bank_complaint_products <- count(bank_complaint_data, "Product")
ranked_bank_complaint_products <- bank_complaint_products %>% arrange(desc(freq))
head(ranked_bank_complaint_products, 30)
```

## The section above in two parts takes the set of complaint types that customers have filed against the banks and then organizes a list of types of products that have had the most issues and the frequency of such complaints. The top three results with many occurances all relate to credit products and specifically with credit reporting being notable. I assumed mortgage products would be highest but this is not the case.  Most issues at the top are either poor investigation quality or incorrect information in reports. Reviewing sub.issue is important to delve into the specifics of what is going on. 
```{r}
bank_complaint_sub_issues <- count(bank_complaint_data, "Sub.issue")
ranked_bank_complaint_sub_issues <- bank_complaint_sub_issues %>% arrange(desc(freq))
head(ranked_bank_complaint_sub_issues, 30)
```

## The subissues here corroborate the data in the prior section as information is either incorrect or status is incorrect.  This supports more reports are filed for data on the consumer rather than a product in its self.  This is against my initial hypothesis that poor products or dishonest selling of goods was responsible.  Now we can look at the banks/institutions of interest for these complaints. 
```{r}
bank_complaint_companies <- count(bank_complaint_data, "Company")
ranked_bank_complaint_companies <- bank_complaint_companies %>% arrange(desc(freq))
head(ranked_bank_complaint_companies, 30)
```

## As expected from the other data, the firms most responsble for issues were the three credit monitoring bureaus by a wide margin versus every other institution.  This is the result of one of two things, there are a significant issue of the credit monitoring bureaus providing good service, and/or these products and services generate a high amount of compalints against other goods. So a mortgage has rarely an issue if conducted properly in documentation, while credit that is incorrectly reported will get the attention of customers.  Credit is important to access credit goods, so customers watch these indicators very closely. Now lets use choropleth data to map how complaints occur across the U.S.
```{r}
data(state.regions)
bank_complaint_choropleth_data <- bank_complaint_data %>% group_by(State) %>% summarize(Frequency = count(State))
bank_complaint_choropleth_data <- bank_complaint_choropleth_data[-c(1,2,3,6,8,16,18,30,34,49,55,58), ]
bank_complaint_choropleth_data <- cbind(bank_complaint_choropleth_data, state.regions)

bank_complaint_choropleth_final <- subset(bank_complaint_choropleth_data, select = -c(1,4,5,6))
colnames(bank_complaint_choropleth_final) <- c("value", "region")

bank_complaint_choropleth_plot <- state_choropleth(bank_complaint_choropleth_final, title  = "US Bank Complaint Reporting by State", legend = "Number of Complaint Reports")
bank_complaint_choropleth_plot
```

## The data in the choropleth map support where the highest number of complaints occur without respect to population per capita.  The data shows that the northeast and California have highest number of complaints filed while there is a notable amount of complaints filed in Georgia, Illinois, Florida, and Texas.  These areas have a marginally high income but also have a high number of urban centers or older populations that can be seen as vulnerable.  
```{r}
complaints_over_time <- bank_complaint_data %>% select(1,18)
complaints_over_time$year_month_day <-mdy(complaints_over_time$Date.received)
complaints_over_time$year_month <- format(complaints_over_time$year_month_day, "%Y-%m")
complaints_over_time$year <- format(complaints_over_time$year_month_day, "%Y")

complaints_time_series_y_m <- complaints_over_time %>% 
  group_by(year_month) %>%
  dplyr::summarize(count = n())

complaints_time_series_y <- complaints_over_time %>% 
  group_by(year) %>%
  dplyr::summarize(count = n())

complaints_time_series_year_plot <- ggplot(data = complaints_time_series_y, aes(year,count, group=1)) +
  geom_line() +
  geom_point() + 
  scale_y_continuous(label=scales::comma) +
  ggtitle("Complaint Count by Year")
complaints_time_series_year_plot
```

## The charting above is the year by year output of complaints since the reporting began in 2011.  There is a notable spike from 2017 to 2020 in the number of complaints filed. Chances are that this spike is due to the increasing reporting requirements from lenders and higher awareness of consumers to file a complaint, rather a dramatic increase in bad reporting or fradulent activity. 2021 is only lower due to the fact that we are only 4 months into the year so its not a fair comparison.  

```{r}
bank_complaint_firms <- bank_complaint_data %>% select(8,18)

bank_complaint_firms$Firm_Type <- NA


bank_complaint_firms$Firm_Type[bank_complaint_firms$Company == "EQUIFAX, INC."] <- "Credit_Bureau"
bank_complaint_firms$Firm_Type[bank_complaint_firms$Company == "TRANSUNION INTERMEDIATE HOLDINGS, INC."] <- "Credit_Bureau"
bank_complaint_firms$Firm_Type[bank_complaint_firms$Company == "Experian Information Solutions Inc."] <- "Credit_Bureau"

bank_complaint_firms[,c("Firm_Type")][is.na(bank_complaint_firms[,c("Firm_Type")])] <- "Bank"
View(bank_complaint_firms)

bank_complaint_firms_counts <- count(bank_complaint_firms, "Firm_Type")

bank_complaint_firms_pareto <- ggplot(bank_complaint_firms_counts, aes(x=Firm_Type, y=freq)) +
 stat_pareto(point.color = "red",
             point.size = 3,
             line.color = "black",
             #size.line = 1,
             bars.fill = c("blue", "red"),
 )
bank_complaint_firms_pareto

firm_frequencies <- freq(bank_complaint_firms$Firm_Type)
firm_frequencies
```
## The last graph of this section supports a Pareto analysis of the proportion of complaints that occur for banking institutions against credit bureaus.  As a whole we see three firms, the credit bureaus, against all other firms.  This is signficant because the banks have thousands of companies listed and have only about half the count of complaints.  The credit bureaus have a disproportionate amount of issues relative to all other institutions.  A pareto analysis is a good graphic measure of this information.  
